CRUSE to Safe Cycling in Ireland

### 2023-10-03 paper.qmd
Add aim

Mobility kills.
High energy transport systems are a major contributor to climate change, a leading cause of premature death and injury due to road traffic collisions, and a cause of disease due to airborne pollutants.
Conversely, evidence-based and effective transport *policies* have great potential to decarbonize economies, improve public health, and save lives.

Transport is responsible for 23% of global greenhouse gas emissions, 70% of which are from road transport, nearly half of which (around 10% of global emissions) is from passenger cars [@jaramillo2022].
The transport system encourages, enables and in some cases enforces unsustainable lifestyles.
Services that are only accessible by car lock-in car dependency [@gray2001; @shergold2012; @motte-baumvol2010].

Recognizing the growing evidence of such impacts of poorly designed and performing transport systems, governments in many countries have set targets and taken actions.
In the context of climate, road safety and physicial inactivity crises, policies to improve transport systems can be classified according to the 'Avoid-Shift-Improve' (ASI) framework [@jaramillo2022].
The framework highlights the importance of demand reduction (*avoid*ing unnecessary trips), in addition to mode *shift* to sustainable modes and *improvement* of existing energy converters, in that order.

@@

A key consideration for TII is to ensure both urban and rural Ireland were targeted for cycling infrastructure, as shown in @fig-dublin.
To this end, TII extensively reviewed activity on Irish Roads and how the road network was being used by commuters, cyclists, heavy goods vehicles and other road users across the Irish transport network.
The aim was for the tool to provide strong, national, systematic but locally-specific evidence to monitor cycling friendliness and safety in Ireland, to ensure strategic alignment with national, regional and local policies.
The open source and publicly available nature of the tool ensures more inclusive and evidence-based conversations around cycle network planning between all stakeholders in the transport planning process.
Other use cases in the Irish context includes monitoring the need for existing cycling infrastructure upgrades, undertaking appraisal for business case development, as well as safety and exposure information.
The latter is in line with the European Cycling Federation, Dutch Cycling Embassy recommendations and the European RISM Directive, for reporting collision rates of vulnerable road users by 2024.
These factors mean that the tool can be seen as an open access 'leverage point' in the planning system [@lovelace2020].

@@

The rest of paper is structured as follows.
In @sec-methods, we outline the methods used to generate the evidence presented in the CRUSE Tool for Ireland.
In @sec-results, we present the results of the CRUSE Tool, including estimates of current cycling levels and future cycling potential at the national, regional and local levels.
In @sec-discussion, we discuss the implications of the results for policy and practice.
In @sec-conclusions, we conclude the paper.

>>>>

<!-- Mobility kills. -->
High energy transport systems are a major contributor to climate change, a leading cause of premature death and injury due to road traffic collisions, and a cause of disease due to airborne and noise pollutants.
Conversely, evidence-based and effective transport *policies* have great potential to decarbonize economies, improve public health, and save lives.

Transport is responsible for 23% of global greenhouse gas emissions, 70% of which are from road transport. Nearly nearly half of these emissions (around 10% of global emissions) are from passenger cars [@jaramillo2022].
The transport system encourages, enables and in some cases enforces unsustainable lifestyles.
Services that are only accessible by car lock-in car dependency [@gray2001; @shergold2012; @motte-baumvol2010].

Recognizing the growing evidence of such impacts of poorly designed and performing transport systems, governments in many countries have set targets and taken actions.
In the context of climate, road safety and physicial inactivity crises, policies to improve transport systems can be classified according to the 'Avoid-Shift-Improve' (ASI) framework [@jaramillo2022].
The ASI framework highlights the importance of demand reduction (*avoid*ing unnecessary trips), in addition to mode *shift* to sustainable modes and *improvement* of existing energy converters, in that order.

@@

A key consideration for TII is to ensure both urban and rural Ireland were targeted for cycling infrastructure, as shown in @fig-dublin.
To this end, TII extensively reviewed activity on Irish Roads and how the road network was being used by commuters, cyclists, heavy goods vehicles and other road users across the Irish transport network.
The aim was for the tool to provide strong, national, systematic but locally-specific evidence to monitor cycling friendliness and safety in Ireland, to ensure strategic alignment with national, regional and local policies.
The open source and publicly available nature of the tool ensures more inclusive and evidence-based conversations around cycle network planning between all stakeholders in the transport planning process.

TII also requires evidence to support monitoring of growth in cycling, upgrades of existing cycling infrastructure, business case development and appraisal, and exposure information to estimate crash rates.
Exposure information is a key part of the the European RISM Directive, for reporting collision rates of vulnerable road users by 2024.
These factors mean that the tool can be seen as an open access 'leverage point', providing key information for and supporting ambitious plans in many aspects the planning system, from network design to post-build monitoring [@lovelace2020].

@@

The main aim of the paper is to present the CRUSE Tool and describe the methods and datasets on which the estimates of current and future potential cycling levels are based.
We will also highlight key results, discuss the implications for policy and practice, and outline how similar tools could be developed in other countries.
It is structured as follows.
In @sec-methods, we outline the methods used to generate the evidence presented in the CRUSE Tool for Ireland.
In @sec-results, we present the results of the CRUSE Tool, including estimates of current cycling levels and future cycling potential at the national, regional and local levels.
In @sec-discussion, we discuss the implications of the results for policy and practice.
In @sec-conclusions, we conclude the paper.
